Conversations in time: interactive visualisation to explore structured temporal data

### 2020-09-13 tsibbletalk.Rmd
abstract

---
title: "Conversations in time: interactive visualisation to explore structured temporal data"
author:
- name: Earo Wang
  affiliation: The University of Auckland
  address: Auckland, New Zealand
  email: earo.wang@auckland.ac.nz
- name: Dianne Cook
  affiliation: Monash University
  address: Melbourne, Australia
  email: dicook@monash.edu
bibliography: tsibbletalk.bib
abstract: >
  An abstract of less than 150 words.
preamble: |
  \usepackage{subfig}
output: rticles::rjournal_article
---

>>>>

---
title: "Conversations in time: interactive visualisation to explore structured temporal data"
author:
- name: Earo Wang
  affiliation: The University of Auckland
  address: Auckland, New Zealand
  email: earo.wang@auckland.ac.nz
- name: Dianne Cook
  affiliation: Monash University
  address: Melbourne, Australia
  email: dicook@monash.edu
bibliography: tsibbletalk.bib
abstract: >
  The \pkg{tsibbletalk} package can be used to interactively explore temporal data
  of nesting and crossing structures, and discover interesting periodic/aperiodic
  slices. This package implements a shared `tsibble` object that allows for
  linked brushing for coordinated views, and a shiny module that aids in wrapping
  time lines for seasonal patterns. The basic usage of functions is demonstrated
  in the article, wit data examples.
preamble: |
  \usepackage{subfig}
output: rticles::rjournal_article
---

### 2020-09-13 tsibbletalk.tex
abstract

\abstract{%
An abstract of less than 150 words.
}

@@

R \CRANpkg{shiny} \citep{R-shiny} and \CRANpkg{htmlwidgets}
\citep{R-htmlwidgets} lay a fundamental infrastructure in marrying R
with HTML elements and Javascript for interactivity, although they do
not emphasise on interactive graphics. The \CRANpkg{htmlwidgets} package
opens a door to embedding Javascript libraries into R such that users
are able to write only R code to generate web-based plots. Many
Javascript charting libraries have been ported to R as HTML widgets,
including \CRANpkg{plotly} \citep{plotly2020}, \CRANpkg{rbokeh}
\citep{R-rbokeh}, and \CRANpkg{leaflet} \citep{R-leaflet} for maps. To
enable interactions between different widgets, it can be achieved with
\CRANpkg{shiny} or \CRANpkg{crosstalk} \citep{R-crosstalk}. The
\CRANpkg{crosstalk} extends \CRANpkg{htmlwidgets} with shared R6
instances to support linked brushing and filtering across widgets,
without relying on \CRANpkg{shiny}.

@@

This shiny module, decoupled to \texttt{tsibbleWrapUI()} and
\texttt{tsibbleWrapServer()}, presents a clean interface and forms a
resusable piece in a shiny application. A shiny module provides the
vehicle in modularising shiny applications for both users and
developers. Like all shiny modules, the first argument in both functions
requires a user-supplied session id that must be unique. The UI function
\texttt{tsibbleWrapUI()} simply shows a slider that animates or controls
the number of periods to be diced. The workhorse is certainly the server
function \texttt{tsibbleWrapServer()}, encapsulating the algorithm that
transforms data and sends messages to update the plot accordingly. The
\texttt{plot} argument expects a \texttt{ggplot} or \texttt{plotly}
object, where one can plot data using either lines or other graphical
elements (such as boxplots). As the function name suggests, a (shared)
tsibble is needed to start the engine, and thereby the time
\texttt{index} can be retrieved for dissection. The \texttt{period}
option semantically takes a desired number of seasonal periods to be
shifted, for example data shifted by ``1 day'', ``2 days'', or ``1
week'', etc. In other words, the \texttt{period} defines the grind
level. As for date-times (represented by \texttt{POSIXt}), it ranges
from from fine ``day'' to coarse ``year''. The following code snippet
generates Figure \ref{fig:wrap-ped}.

@@

ui <- fluidPage(
  tsibbleDiceUI("dice")
)
server <- function(input, output, session) {
  tsibbleDiceServer("dice", ggplotly(p_line, height = 700), period = "1 day")
}
shinyApp(ui, server)
\end{Sinput}
\end{Schunk}

>>>>

\abstract{%
The \pkg{tsibbletalk} package can be used to interactively explore
temporal data of nesting and crossing structures, and discover
interesting periodic/aperiodic slices. This package implements a shared
\texttt{tsibble} object that allows for linked brushing for coordinated
views, and a shiny module that aids in wrapping time lines for seasonal
patterns. The basic usage of functions is demonstrated in the article,
wit data examples.
}

@@

R \CRANpkg{shiny} \citep{R-shiny} and \CRANpkg{htmlwidgets}
\citep{R-htmlwidgets} lay a fundamental infrastructure in marrying R
with HTML elements and Javascript for interactivity, although they
seldom emphasise on interactive graphics. The \CRANpkg{htmlwidgets}
package opens a door to embedding Javascript libraries into R such that
users are able to write only R code to generate web-based plots. Many
Javascript charting libraries have been ported to R as HTML widgets,
including \CRANpkg{plotly} \citep{plotly2020}, \CRANpkg{rbokeh}
\citep{R-rbokeh}, and \CRANpkg{leaflet} \citep{R-leaflet} for maps. To
enable interactions between different widgets, it can be achieved with
\CRANpkg{shiny} or \CRANpkg{crosstalk} \citep{R-crosstalk}. The
\CRANpkg{crosstalk} extends \CRANpkg{htmlwidgets} with shared R6
instances to support linked brushing and filtering across widgets,
without relying on \CRANpkg{shiny}.

@@

This shiny module, decoupled to \texttt{tsibbleWrapUI()} and
\texttt{tsibbleWrapServer()}, presents a clean interface and forms a
resusable piece in a shiny application. A shiny module provides the
vehicle in modularising shiny applications for both users and
developers. Like all shiny modules, the first argument in both functions
requires a user-supplied session id that must be unique. The UI function
\texttt{tsibbleWrapUI()} simply shows a slider that animates or controls
the number of periods to be diced. The workhorse is certainly the server
function \texttt{tsibbleWrapServer()}, encapsulating the algorithm that
transforms data and sends messages to update the plot accordingly. The
\texttt{plot} argument expects a \texttt{ggplot} or \texttt{plotly}
object, where one can plot data using either lines or other graphical
elements (such as boxplots). As the function name suggests, a (shared)
tsibble is needed to start the engine, and thereby the time
\texttt{index} can be retrieved for dissection. The \texttt{period}
option semantically takes a desired number of seasonal periods to be
shifted, for example data shifted by ``1 day'', ``2 days'', or ``1
week'', etc. In other words, the \texttt{period} defines the grind
level. For date-times (represented by \texttt{POSIXt}), the adjustment
ranges from from fine ``day'' to coarse ``year''. The following code
snippet generates Figure \ref{fig:wrap-ped}.

@@

ui <- fluidPage(
  tsibbleWrapUI("dice")
)
server <- function(input, output, session) {
  tsibbleWrapServer("dice", ggplotly(p_line, height = 700), period = "1 day")
}
shinyApp(ui, server)
\end{Sinput}
\end{Schunk}
